---
title: "Colombia Health Loss"
output: html_document
---

In this exercise, we'll compare how different **metrics of disease burden** tell different stories about the health of a country. In particular, we'll be using data from the Colombia (see `README.md` for more information). 

```{r, message=FALSE, warning=FALSE, echo=FALSE}
# Set up 
library(dplyr)
library(knitr)
setwd('~/Documents/info-498c/demos/gbd/dalys')


# Load data
colombia.data <- read.csv('data/prepped/overview.csv')

```

## Deaths
We'll begin with a simple metric of disease burden, _number of deaths_. Using that metric, here are the 10 diseases that cause the most health loss in Colombia:

```{r, echo=FALSE}
# Begin by calculating two new columns:
# - burden.rank: rank *within metric* for each disease
# - percent: a rounded percentage of the "Value" column
ranked.data <- colombia.data %>% 
                  group_by(metric) %>% 
                  mutate(burden.rank = rank(desc(Value)),
                       percent = paste0(round(Value * 100, 1), '%')
                       ) %>% 
                  ungroup()
                  
# Create a summary table of the top 10 burdensome diseases as measured by deaths


# Use the `kable` function to show the table you've created

```

## YLLs
As a more nuanced metric, we'll consider _Years of Life Lost_. This measure incorporates the age at death to apply a higher burden to diseases that affect younger people. Below is a bar chart of the top 10 diseases as measured by **YLLs**. 

```{r, echo=FALSE}
# Set margin spacing for graphics
par(mar=c(5,10,4,2))



# Filter down to top YLLs



# Create a barplot of the top 10 burdensome diseases as measured by YLLs




```

## YLDs
We can perform a similar analysis of Years Lived with Disability (YLDs).

```{r, echo=FALSE}
# Create a summary table of the top 10 burdensome diseases as measured by YLDs



# Use the `kable` function to show the table you've created


```

## DALYs
Finally, we can see this breakdown for DALYs:

```{r, echo=FALSE}
# Create a summary table of the top 10 burdensome diseases as measured by DALYs

# Use the `kable` function to show the table you've created


```


```{r, echo=FALSE}
# Bonus: Create an arrow diagram (http://ihmeuw.org/3zn7) comparing deaths to YLLs
```